Actuator line modeling of vertical-axis turbines

### 2016-05-13 paper.tex
Combine turbines section with last intro paragraph

The ASSM and ALM were implemented to model a very low Reynolds number 2-D
cross-flow turbine experiment in a flume using large eddy simulation
(LES)~\cite{Shamsoddin2014}. Performance predictions for this case were not
reported, but the ALM was shown to be more effective at postdicting the wake
characteristics measured in the experiments by Brochier et
al.~\cite{Brochier1986}.

It is therefore proposed that an actuator line model may be the optimal
combination of high-fidelity flow modeling that includes performance
predictions, but with reduced computational expense. Here we have developed an
ALM for cross-flow turbines inside both RANS and LES simulations, implemented as
an extension for the OpenFOAM open-source CFD library. This model was validated
against experimental data for both a high \cite{Bachant2015-JoT} and a medium
\cite{Bachant2016-RM2-paper} solidity vertical-axis turbine.

\section{Turbines}

Two turbines were modeled: the high solidity ($c/R=0.28$) UNH Reference
Vertical-Axis Turbine (UNH-RVAT) and the medium solidity ($c/R=0.07$--$0.12$) US
Department of Energy/Sandia National Labs Reference Model 2 (RM2) cross-flow
turbine (at 1:6 scale), both of which are shown in Figure~\ref{fig:turbines}.
Both rotors are of similar geometric size ($\sim 1$ m) and are constructed from
symmetrical NACA foils mounted at mid-chord and mid-span.

Open datasets for the performance and near-wake dynamics of the UNH-RVAT and
RM2, available from \cite{Bachant2014-RVAT-baseline,Bachant2016-RVAT-Re-dep} and
\cite{Bachant2016-RM2-data}, respectively. have been acquired in the UNH towing
tank at Reynolds numbers based on turbine diameter $Re_D = 10^6$. Analysis of
these datasets showed that turbine performance and near-wake characteristics
were nearly Reynolds number independent in this
regime~\cite{Bachant2016-Energies,Bachant2016-RM2-paper}.

>>>>

The ASSM and ALM were implemented to model a very low Reynolds number (based on
rotor diameter, $Re_D \sim 10^4$) 2-D cross-flow turbine experiment in a flume
using large eddy simulation (LES)~\cite{Shamsoddin2014}. Performance predictions
for this case were not reported, but the ALM was shown to be more effective at
postdicting the wake characteristics measured in the experiments by Brochier et
al.~\cite{Brochier1986}.

It is therefore proposed that an actuator line model may be the optimal
combination of high-fidelity flow modeling that includes performance
predictions, but with reduced computational expense. Here we have developed an
ALM for cross-flow turbines inside both RANS and LES simulations, implemented as
an extension for the OpenFOAM open-source CFD library. This model was validated
against experimental data for the high solidity ($c/R=0.28$) UNH Reference
Vertical-Axis Turbine (UNH-RVAT) and the medium solidity ($c/R=0.07$--$0.12$) US
Department of Energy/Sandia National Labs Reference Model 2 (RM2) cross-flow
turbine (at 1:6 scale), both of which are shown in Figure~\ref{fig:turbines}.
Validation datasets for both turbines were taken from
\cite{Bachant2016-RVAT-Re-dep} and \cite{Bachant2016-RM2-data}, respectively.
The turbines were modeled at $Re_D \sim 10^6$, at which experimentally measured
performance and near-wake characteristics were nearly Reynolds number
independent~\cite{Bachant2016-Energies,Bachant2016-RM2-paper}.
